---
title: "SimpleDynamicModels"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
```

Dynammic models always involves derivatives (equations that express how things change from time step to time step or place to place )

# Derivates 

Implement population growth as a derivative - a model of population change

```{r derivaties, echo=FALSE}

# note that we include time here but we don't use it; we will need this later
source("./dexppop.R")

# see how it works
dexppop(P=20, r=0.01)


#what is this?

# notices this is the same as
dexppop(t=100,P=20, r=0.01)

# lets look at this for a range of initial populations
pops = seq(from=1, to=100)
tmp = pops %>% map(~dexppop( time=0,r=0.01, P=.x))
pchange = unlist(tmp)


pdyn = data.frame(pops, pchange)
ggplot(pdyn, aes(pops, pchange))+geom_point(col="green", size=1.5)


# why is this a straight line?
# how many new individuals are born at each population level

# try this - add a carrying capacity (dP/dt = 0 if P > carryingcapacity)
```

```{r}
dexppop_play.2 = function(time, P, r, K) {
  
  # compute rate of change of population
  dexpop = r*P
  
  # set rate of change to 0 if P is greater than carrying capacity
  dexpop = ifelse(P > K, 0, dexpop)
  return(list(dexpop))
}

pops = seq(from=1, to=100)
tmp = pops %>% map(~dexppop_play.2( time=0,r=0.01,K=75, P=.x))
pchange = unlist(tmp)


pdyn = data.frame(pops, pchange)
ggplot(pdyn, aes(pops, pchange))+geom_point(col="green", size=1.5)


```


# Integration

What if we wanted to look at how population grows over time - population in 20 years given an initial condition



Some code to look at simple population growth models as
dynamic models where
* explicit solution to differential equation is known; e.g. you can integrate both sides of the equation! Not always possible but lets look at a case where it is possible
* must be solved by iteration; this is what we do when we can't integrate both sides 

# Explicit Solution is available

```{r explicit}
source("./exppop.R")

# gives population after any time given an initial population

# 20 rabbits, growth rate of 0.01 how many in 30 years
exppop(T=30, P0=20, r=0.01, K=1000)

# if we want to see how populatin evolves over time - generate a time series 

initialrabbits = 20
years = seq(from=1, to=100, by=2)
Ptime = years %>% map_dbl(~exppop( P0=initialrabbits, r=0.01, K=1000, T=.x))

# keep track of what times we ran
Ptime = data.frame(P=Ptime, years=years)

ggplot(Ptime, aes(years,P))+geom_point()+labs(x="years",y="Rabbit Population")

# try generating results for maximum and minimum possible r values to compare (guess at what you think)
max_r = 0.1
min_r = 0.01
K = 1000

tmp = years  %>% map_dbl(~exppop(r=max_r, P0=initialrabbits, K=K, T=.x))
Ptime$Pmaxr = tmp
tmp = years %>% map_dbl(~exppop(r=min_r, P0=initialrabbits, K=K, T=.x))
Ptime$Pminr = tmp

head(Ptime)
Ptimep = Ptime %>% gather(key="r",value="P",-years)
ggplot(Ptimep, aes(years,P, col=r))+geom_point()+labs(x="years",y="Rabbit Population")

# notice how populatin becomes unstable for high growth rates! 


```

Using a solver....when you can't do the integration by hand :)

For example, if you added a carrying capacity!

We can still do the integration by iteration - approximates it
Implement the differential equation as a function that

*returns the derivative (as a list)
*inputs time, the variable(s) and a parameter list

(it needs time even though you don’t use it)

I name them starting with *d* to remind myself that they
are computing a derivative


R has a solver called *ODE* for solving ordinary differential equations
from package **desolve**

*ODE* requires

* initial conditions
* values of independent where we want values of dependent variable (e.g times where we want population)
* the derivative as a function
* a list that contains alll  parameter values (or if you have only one parameter then you can use a single value)

```{r usingode}


source("./dexppop.R")
library(deSolve)
initialrabbits = 20
years = seq(from=1, to=100, by=2)

# run the solver
Ptime = ode(y=initialrabbits, times=years, func=dexppop,parms=c(0.01))
head(Ptime)
colnames(Ptime)=c("year","P")

# notice that there are additional pieces of information year, including the method used for integration
attributes(Ptime)

# this also means you need to extract just the data frame for plotting
ggplot(as.data.frame(Ptime), aes(year,P))+geom_point()+labs(y="Population", "years")

# this also works (of course function can be by order)
Ptime=ode(initialrabbits, years, dexppop,0.03)
colnames(Ptime)=c("year","P")
ggplot(as.data.frame(Ptime), aes(year,P))+geom_point()+labs(y="Population", "years")

# how would we add carrying capacity here

```


You can play a bit with changing your function to something that you can't integrate "by hand"

BUT we might want more parameters

to work with ODE,
parmeters must all be input as a single list; simlar to how we return multiple outputs from a function

see example below..lets add a carrying capacity


```{r}



source("./dexppop_play.R")

# create parameter list
initalrabbits=2
newparms = list(r=0.03, carry_capacity=300)

#apply solver
results=ode(initialrabbits, years, dexppop_play,newparms)
head(results)

# add more meaningful names
colnames(results)=c("year","P")

# plot
ggplot(as.data.frame(results), aes(year,P))+geom_point()+labs(y="Population", "years")


# try again with different parameters
alternativeparms = list(r=0.04, carry_capacity=500)
results2=ode(initialrabbits, years, dexppop_play,alternativeparms)


# look at results
head(results2)
colnames(results2)=c("year","P_parm2")

# plot
ggplot(as.data.frame(results2), aes(year,P_parm2))+geom_point()+labs(y="Population", "years")

# compare by combining into a single data frame
both = inner_join(as.data.frame(results), as.data.frame(results2))

both_p = both %>% gather(key="model", value="P", -year)
ggplot(both_p, aes(year,P, col=model))+geom_point()+labs(y="Population", "years")

# try playing on your own - modiffy the function in some way
```

