---
title: "Diffusion"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Diffusion

Diffusion can be implemented as a partial differential equation 
Complicated to solve - but there are tool in R for specific types of partial differential equations
![Reactive Transport Example](https://cran.r-project.org/web/packages/ReacTran/index.html)

More info on differential equations in R
![online book differential equation in R](http://www-labs.iro.umontreal.ca/~mignotte/IFT2425/Documents/Solving_Differential_Equations_In_R_Soetaert_K.pdf)

But we can appoximate this with a difference equation - and iterative to get an estimate of how diffusion works (see class lecture pdf)


Now for partial derivatives - time and space! it gets much more tricky ...beyond this course 

Example of Diffusion - difference equation implementation to see what some issues can be


```{r}

source("./diffusion.R")

# run our diffusion model (iterative difference equation) with initial concentration of 10, for 8 timestep (size 1m), and 10 space steps (size 1s)
# using diffusion parameters 0.5 s/m2, 10 m2
result = diff1(initialC=10, nx=10, dx=1, nt=8, dt=1, D=0.5, area=10)

# a list is returned with our 3 data frames for concentration (conc), qin and qout
result

# used filled contour to plot results
head(result$conc)
filled.contour(result$conc, xlab="Time", ylab="Distance")

# or if you prefer this orientation (Distance on x axis)
filled.contour(t(result$conc), ylab="Time", xlab="Distance")

# changes diffusivity and other parameters particularly
# diffusivity, dx and dt
res=diff1(initialC=10,nx=10,dx=1,nt=10,dt=30,D=0.006,area=1)

filled.contour(res$conc, xlab="Time", ylab="Distance")


# we can also see how much material moved from place to place each time step
filled.contour(res$qin, xlab="Time", ylab="Distance")

# what if we increase diffusivity
resfast=diff1(initialC=10,nx=10,dx=1,nt=10,dt=10,D=0.08,area=1)
filled.contour(resfast$conc, xlab="Time", ylab="Distance")


# play with time step, space step and parameters


```

```{r}

# Discretization Issue Example
resunstable=diff1(initialC=10,nx=10,dx=1,nt=10,dt=10,D=0.8,area=1)
filled.contour(resunstable$conc, xlab="time",ylab="Distance Along Path")

# this illustrates the problem with difference equations (and the challenges that methods for numerical integration try to overcome)
# if things are changing quickly we need to use much smaller time, space steps to avoid overshoot and instability

# so lets cut our step size by 10 (dt) (but then add 10 more steps (nx) to cover the same distance)
resunstable=diff1(initialC=10,nx=100,dx=1,nt=10,dt=1,D=0.8,area=1)
filled.contour(resunstable$conc, xlab="time",ylab="Distance Along Path")

```

