---
title: "ESM 232 Assignment 5"
author: "Ruben Sanchez Ramirez"
date: "April 26, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}

# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(ggplot2)
library(purrr)
library(sensitivity)
library(pse)

```

Dynamic Models - Assignment 5

In Class, we learned how to use the ODE solver to generate results from a dynamic model (where the value of something is evolving over time and this evolution is described by a differential equation). We started with dP/dT = rP (a simple model of population growth).  We then looked at adding a carrying capacity dP/dT = rP until P is at carrying capacity and then dP/dT = 0.  

Another way to model carrying capacity is to assume that dP/dT = r*P*(1-P/K)  where K is carrying capacity.  Implement this differential equation as a function.  Then run the model using the ODE solver, with the following 

r = 0.05

K = 20

initial P is 1

graph results for time from 1 to 50 years.Submit the graph, your Rmarkdown on Gauchospace

If you want to check your work - the population at year 50 should be 7.576973
