---
title: "using_functions"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Organizing your work

* Store function definition in **R** subdirectory
* Store  Rmarkdown in different subdirectory (**Rmarkdown**)
* Set the scripts (**Rmarkdown**) as working directory
* Use **source** to read the functions into your working directory

```{r getfunctions}
# you may need to change the name here if you save power_gen.R in a different place
source("../R/power_gen.R")
power_gen
```

```{r applyfunctions}

# parameters are ordered
power_gen(20,1)
# uses 20 for height and 1 for flow
power_gen(height=20, flow=1)

# defaults are used, unless specified
# consider a less efficient turbine
power_gen(20,1, Keff=0.5)

#how would you calculate power generation on Mars where
# gravity is 10 time lighter

# how about a high efficieny turbine with denser water (0.9 for efficiency; denser water 1200)


# you can store the results of a function in a variable
result_low = power_gen(20,1,Keff=0.2)
result_high = power_gen(20,1,Keff=0.8)

# now compute the difference in power between a high and low efficient turbine
result_high-result_low
```
## Warning (a good to know piece of info)

Note that by *sourcing* a function - it will essentially overwrite anything else in your workspace with the same name 

*Warning* - if you name your function, the same as an internal R function, your new function will take precidence, and **hide** the internal R function

In R, functions are organized in **Packages**
You've probably loaded different packages, that provide different functions
There are a number of packages **base**, **stats** that are automatically loaded
You can usually find the package associated with any function from online help

* consider **runif** function in the **stats** package

[runif help][https://stat.ethz.ch/R-manual/R-patched/library/stats/html/Uniform.html]

To use a function associated with a particular package

**package::function**

This is helpful if you end up using functions with the same name as functions in other packages

```{r  functionnaming}

# consider runif an internal R function that returns 'n' random numbers between a minimum and maximum value
runif(min=0, max=10, n=3)

#what if I create my own function called runif

runif = function(x,y) {
  result=x**y
  return(result)
}
#runif(min=0, max=10, n=3)

runif(x=2,y=3)

stats::runif(min=0,max=10, n=3)

# if your remove your runif it will default back to core package 
rm(runif)
#runif(x=2,y=3)
runif(min=0,max=10, n=3)
```



