---
title: "Untitled"
author: "Ruben Sanchez Ramirez"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Read in data 
clim <- read_table2(here::here("assignments","data","clim.txt")) %>% 
  clean_names()
```

```{r first assignment}
###Naomi's almond function
  #Changes: 
    #Jan_P month== was 2, changed to 1
    #Function output statistics instead of yeild; changed to output matrix of yield
almond = function(clim,Tmincoeff1=-0.015, Tmincoeff2=-0.0046, Pcoeff1=-0.07, Pcoeff2=0.0043, intercep=0.28) {
# find need climate variables
tmp = clim %>% group_by(month, year) %>% dplyr::summarize(tmin_c=mean(tmin_c))
Feb_minT = (tmp %>% subset(month==2))$tmin_c 
tmp = clim %>% group_by(month, year) %>% dplyr::summarize(precip=sum(precip))
Jan_P = (tmp %>% subset(month==1))$precip
#compute yield
yield =
  Tmincoeff1*Feb_minT+Tmincoeff2*Feb_minT**2 +
  Pcoeff1*Jan_P + Pcoeff2*Jan_P**2 + intercep
return(yield)
}
##calculate yield
ANNyield <- cbind(unique(clim$year), almond(clim))
colnames(ANNyield) <- c("year", "yield")
ANNyield <- as.data.frame(ANNyield)
#plot
ggplot(ANNyield, aes(year, yield)) + geom_bar(stat="identity")
```



```{r second assignment}
###Had to change function again so that climate is not subset inside almond function
tmp = clim %>% group_by(month, year) %>% dplyr::summarize(tmin_c=mean(tmin_c))
Feb_minT = (tmp %>% subset(month==2))$tmin_c 
tmp = clim %>% group_by(month, year) %>% dplyr::summarize(precip=sum(precip))
Jan_P = (tmp %>% subset(month==1))$precip
##Changed almong function to just output vector of yield
almond = function(T, P,Tmincoeff1=-0.015, Tmincoeff2=-0.0046, Pcoeff1=-0.07, Pcoeff2=0.0043, intercep=0.28) {
#compute yield
yield =
  Tmincoeff1*T+Tmincoeff2*T**2 +
  Pcoeff1*P + Pcoeff2*P**2 + intercep
  #compute statistics
return(yield)
}
#Create input for parameter uncertainty
Pcoeff2_uncert <- rnorm(500,0.0043,0.001)
summary(Pcoeff2_uncert)
#Run almond function across all years and all parameter uncertainty
Psens <- sapply(Pcoeff2_uncert, function(x) {tmp = mapply(FUN=almond, T=Feb_minT, P=Jan_P, Pcoeff2= x); return(tmp)})
#Set up data frame to be plotted
colnames(Psens) = Pcoeff2_uncert
rownames(Psens) = unique(clim$year)[2:23]
unique(clim$year)[2:23]
Psensg = t(Psens) 
Psensg = as.data.frame(Psensg) %>% gather(year, yield)
#plot sensitivity analysis
ggplot(Psensg, aes(year, yield)) + geom_boxplot()
 
```



```{r third assignment}
#Compute_NPV function
compute_NPV = function(value, time, discount=0.12) {
	result = value / (1 + discount)**time
	return(result)
}
#Create data frame to run NPV function over
nruns= length(unique(clim$year))
Alm_yield <- as.data.frame(matrix(nrow=nruns, ncol=4))
colnames(Alm_yield) = c("year","Feb_minT", "Jan_P","anom") 
Alm_yield$year = unique(clim$year) 
Alm_yield$Feb_minT = Feb_minT
Alm_yield$Jan_P = Jan_P
Alm_yield$anom = almond(T = Alm_yield$Feb_minT, P = Alm_yield$Jan_P)
# Almond cost $3800/acre
# Almond value is $2.50/lb
# Almond production is 1/ton/yr/acre
#function output is in tons/acre
	#2000lb/acre/yr * $2.50/lb = $5000/acre/yr on average
#Calculate NPV
Alm_yield$profit = ((1+Alm_yield$anom) * 5000) - 3800
yrs = seq(from=1, to=nruns)
Alm_yield$NPV =  compute_NPV(Alm_yield$profit, time=yrs)
#Sum NPV
sum(Alm_yield$NPV)
